Analysis and Figures for Gallego et al, 2020

### 2020-09-03 Manuscript/What.lies.ahead.reviewed.tex
Change in caption Fig2. Removed sd from calculations. Change Fig 3 axis labels. Added Rmd for model test

\subsection*{eDNA sequencing and bioinformatic processes}
We purified DNA from each filtered sample using a Phenol- Chloroform-Isoamyl Alcohol protocol, following \cite{renshaw2015room}.  After reducing inhibition via a 1/10 to 1/100 dilution, the extract was used as template for a PCR reaction targeting a 313bp fragment of cytochrome oxidase I \cite{leray_new_2013}. PCR reactions were performed in triplicate and sequenced individually to quantify the stochasticity of PCR reactions on a mixed template sample, and we attached secondary indexing tags using a two-step PCR process \cite{odonnellPrimers}. PCR conditions and protocols for sample identification followed \cite{kelly2018tides}, and batches of 49 to 178 multiplexed samples were sequenced using MiSeq v2-500 or v3-600 sequencing kits using manufacturer protocols. On each sequencing run, we added triplicate samples consisting on DNA obtained from species not present in the marine environment under study (Red Kangaroo (\textit{Macropus rufus}) and Ostrich (\textit{Struthio camelus})) to establish quality controls of sample assignment and to quantify levels of 'tag-jumping' or sample-cross-talk \cite{schnell2015tag}.

@@

By treating amplification efficiency as consistent within a given taxon, we created an index of abundance for each taxon across space and time \cite{kelly2019understanding}, using pooled data from technical replicates and mean proportions across biological replicates. {\color{red} RPK: which analyses use the index, rather than presence-absence? make that clear here.}

@@

\begin{figure}%[tbhp]
\centering
\includegraphics[width=1\linewidth]{	Figures_for_MS/Figure2.2020-08-27.png}
\caption{\footnotesize{Biological communities and their relationship with environmental variables in the summer months of the Hood Canal. \textbf{A} Constrained Analysis of Principal Coordinates (CAP) of Bray-Curtis dissimilarities among biological communities, as constrained by pH, temperature and salinity (arrows). \textbf{B} Most-likely cluster as a function of temperature and pH given a multinomial logistic model. \textbf{C} Relative abundance (eDNA index; see \cite{kelly2019understanding}) of the taxa best distinguishing the three communities illustrated (SIMPER analysis). See Supporting Information for full analysis.}}
\label{fig:Fig2}
\end{figure}
%Move this to the methods%
 To explore the suitability of different environmental conditions for each taxon, we modeled the likelihood of taxon presence as a function of area and season (random effects) and temperature and pH (fixed effects).  Salinity is predicted to remain largely unchanged in future scenarios \cite{khangaonkar2019salish}. Using predicted conditions for Puget Sound for the year 2095 \cite{khangaonkar2019salish}, we then projected the probability of presence of each taxon across these conditions over 100 draws of the posterior probabilities of all taxa. We used these draws to 1) show the change in the probability of presence of each taxon for each scenario (i.e. 2017 vs 2095, Fig.3 top panel); 2) calculate the  taxon richness within taxonomic groups across the pH-temperature continuum (Fig. 3 center panel),% 2) show the consequent variation in richness for higher taxa across the pH - temperature continuum (Figure 3B),
  and 3) visualized the change in suitability for each taxon in our dataset individually (selected taxa in Fig. 3 bottom panel).

>>>>

\subsection*{eDNA sequencing and bioinformatic processes}
We purified DNA from each filtered sample using a Phenol-Chloroform-Isoamyl Alcohol protocol, following \cite{renshaw2015room}.  After reducing inhibition via a 1/10 to 1/100 dilution, the extract was used as template for a PCR reaction targeting a 313bp fragment of cytochrome oxidase I \cite{leray_new_2013}. PCR reactions were performed in triplicate and sequenced individually to quantify the stochasticity of PCR reactions on a mixed template sample, and we attached secondary indexing tags using a two-step PCR process \cite{odonnellPrimers}. PCR conditions and protocols for sample identification followed \cite{kelly2018tides}, and batches of 49 to 178 multiplexed samples were sequenced using MiSeq v2-500 or v3-600 sequencing kits using manufacturer protocols. On each sequencing run, we added triplicate samples consisting on DNA obtained from species not present in the marine environment under study (Red Kangaroo (\textit{Macropus rufus}) and Ostrich (\textit{Struthio camelus})) to establish quality controls of sample assignment and to quantify levels of 'tag-jumping' or sample-cross-talk \cite{schnell2015tag}.

@@

By treating amplification efficiency as consistent within a given taxon, we created an index of abundance for each taxon across space and time \cite{kelly2019understanding}, using pooled data from technical replicates and mean proportions across biological replicates. {\color{red} RGS: We used this index of abundance to calculate the multivariate community analysis, while species responses used the presence/absence dataset.  RPK: which analyses use the index, rather than presence-absence? make that clear here.}

@@

\begin{figure}%[tbhp]
\centering
\includegraphics[width=1\linewidth]{	Figures_for_MS/Figure2.2020-08-27.png}
\caption{\footnotesize{Biological communities and their relationship with environmental variables in the summer months of the Hood Canal. \textbf{A} Constrained Analysis of Principal Coordinates (CAP) of Bray-Curtis dissimilarities among biological communities, as constrained by pH, temperature and salinity (arrows). \textbf{B} Most-likely cluster as a function of temperature and pH given a multinomial logistic model {\color{red}(salinity was not represented as it was tightly correlated with temperature)}. \textbf{C} Relative abundance (eDNA index; see \cite{kelly2019understanding}) of the taxa best distinguishing the three communities illustrated (SIMPER analysis). See Supporting Information for full analysis.}}
\label{fig:Fig2}
\end{figure}
%Move this to the methods%
 To explore the suitability of different environmental conditions for each taxon, we modeled the likelihood of taxon presence as a function of area and season (random effects) and temperature and pH (fixed effects).  Salinity is predicted to remain largely unchanged in future scenarios \cite{khangaonkar2019salish}. Using predicted conditions for Puget Sound for the year 2095 \cite{khangaonkar2019salish}, we then projected the probability of presence of each taxon across these conditions over 100 draws of the posterior probabilities of all taxa. We used these draws to 1) show the change in the probability of presence of each taxon for each scenario (i.e. 2017 vs 2095, Fig.3 top panel); 2) calculate the  taxon richness within taxonomic groups across the pH-temperature continuum (Fig. 3 center panel),% 2) show the consequent variation in richness for higher taxa across the pH - temperature continuum (Figure 3B),
  and 3) visualized the change in suitability for each taxon in our dataset individually (selected taxa in Fig. 3 bottom panel).

### 2020-09-03 Scripts/Manuscript.Main.Analysis.Rmd
Change in caption Fig2. Removed sd from calculations. Change Fig 3 axis labels. Added Rmd for model test

data.for.richness %>% 
  filter (phylum %in% c("Bacillariophyta", "Dinophyceae")) %>% 
  ggplot(aes(x = Site, y= Richness )) +
  geom_boxplot(aes(fill = Area)) +
  facet_wrap(~phylum, nrow = 2) +
  labs (y = "") -> right.rich.plot

@@

@@

expanded.future$sd <- apply(posterior.predict.all.taxa,2,sd )
```

@@

phylum.stats %>% 
  nest(-phylum) %>% 
  mutate( plot.by.phyl = map2 (data, phylum, ~ .x %>%
  ggplot(aes()) +
  scale_fill_brewer(name = "Year",  type = "qual", palette = 4, direction = 1,
                    breaks=c("2017","2095")) +
  labs (x = "Standardized richness",
        y = "") +
  facet_wrap(~Area) +
  theme_minimal() +
  theme (strip.text.y = element_text(angle = 0),
        axis.text.y = element_text(face = "italic"), 
        #plot.background   = element_rect(fill = "#E5E9F0"),
        legend.direction = "horizontal",
        legend.position  = "bottom",
        plot.title.position = "plot",
        plot.title = element_text(face = "bold",vjust =  -5 ),
        axis.title.x.top  = element_text(face = "bold",vjust = 3) ) +
        geom_histogram(aes( x= present,after_stat(ncount),
                            fill = as.factor(Year),
                            group= as.factor(Year)),
                       binwidth = 0.025,
                       colour = "black",
                       position = "identity", 
                       alpha = 0.75) +
        scale_x_continuous(position = "top", limits = c(0,1)))) %>% 
  select(-data)-> summary.plot

@@

full.grid %>% 
  group_by(Area, Temperature = TemperatureCelsius, pH = pH_units, phylum) %>% 
  summarise(richness = sum(present),
            sd       = sum(sd^2)) %>% 
#  left_join(richness.by.phylum) %>% 
  group_by(phylum) %>% 
  mutate(st.richness = richness/max(richness)) -> richness.by.phylum
```

@@

# Assembling Figure 3

>>>>

data.for.richness %>% 
  filter (phylum %in% c("Bacillariophyta", "Dinophyceae")) %>% 
  ggplot(aes(x = Site, y= Richness )) +
  geom_boxplot(aes(fill = Area)) +
  facet_wrap(~phylum, nrow = 2, scales = "free_y") +
  labs (y = "") -> right.rich.plot

@@

We start with a global dataset PERMANOVA, to show that the communities from each region are different to each other and require independent analyisis

```{r Global PERMANOVA}
data.plankton %>% 
  separate (event, into = c("Site", "Date"), sep = "_") %>% 
  mutate(Area = case_when(Site %in% c("CP","LK", "FH") ~ "San Juan Island",
                          TRUE                         ~ "Hood Canal")) %>% 
   unite (Site, Date, col = "event", sep = "_") %>% 
  mutate(GLOBAL = "Global") %>% 
  nest(-GLOBAL) %>% 
  mutate(comm.matrix = map (data, ~ tibble.to.table(.x, taxon = taxa, Abundance = Normalized.reads, sample.name = event)),
         bc = map(data, ~tibble_to_matrix(.x, taxon = taxa,Abundance = Normalized.reads, sample.name = event, distance = "bray",transformation = "" )),
         env = map (data, ~tibble_to_env(.x, taxon = taxa, Abundance = Normalized.reads, sample.name = event, everything()) %>% 
                       select(-contains("%")) %>% 
                       separate(event, into = c("Site", "Date"), sep = "_")),
         PERMANOVA = map2(bc, env, function(.x, .y){
           adonis(formula = .x ~ Area,data = .y, permutations = 9999
                    )

         })) %>% pull(PERMANOVA)
```

@@

@@

phylum.stats %>% 
  nest(-phylum) %>% 
  mutate( plot.by.phyl = map2 (data, phylum, ~ .x %>%
  ggplot(aes()) +
  scale_fill_brewer(name = "Year",  type = "qual", palette = 4, direction = 1,
                    breaks=c("2017","2095")) +
  labs (x = "Phylum-wide standardized richness",
        y = "") +
  facet_wrap(~Area) +
  theme_minimal() +
  theme (strip.text.y = element_text(angle = 0),
        axis.text.y = element_text(face = "italic"), 
        #plot.background   = element_rect(fill = "#E5E9F0"),
        legend.direction = "horizontal",
        legend.position  = "bottom",
        plot.title.position = "plot",
        plot.title = element_text(face = "bold",vjust =  -5 ),
        axis.title.x.top  = element_text(face = "bold",vjust = 3) ) +
        geom_histogram(aes( x= present,after_stat(ncount),
                            fill = as.factor(Year),
                            group= as.factor(Year)),
                       binwidth = 0.025,
                       colour = "black",
                       position = "identity", 
                       alpha = 0.75) +
        scale_x_continuous(position = "top", limits = c(0,1)))) %>% 
  select(-data)-> summary.plot

@@

full.grid %>% 
  group_by(Area, Temperature = TemperatureCelsius, pH = pH_units, phylum) %>% 
  summarise(richness = sum(present)) %>% 
#  left_join(richness.by.phylum) %>% 
  group_by(phylum) %>% 
  mutate(st.richness = richness/max(richness)) -> richness.by.phylum
```

@@

### Assembling Figure 3
